Constraining gravitational wave amplitude birefringence with GWTC-3

### 2022-08-11 src/tex/ms.tex
New Content & Graph

GW amplitude birefringence is a property of space which consists of the enhancement of one GW polarisation over the other during the propagation of the wave.
The longer the GW travels, the larger the effect of birefringence.
To illustrate the effect of amplitude birefringence,
the observed amplitudes of two polarisations of an artificial GW signal example are shown in figure \textbf{[X]}.
\textbf{[add figure similar to the animation in the website]}
With birefringence, the amplitude is modified during the propagation,
and the observed amplitudes will be different compare to the amplitudes when they are generated at the binary.

@@

To constrain $\kappa$, we perform parameter estimation (PE) with data from the third LIGO-Virgo catalog \citep{LIGO_2021}, GWTC-3, using Bilby,
a Bayesian toolkit for GW data analysis which is able to calculate posteriors of GW parameters based on interferometer data
and priors for the parameters in question. \citep{Ashton_2019}

@@

@@

\begin{figure}[h!]
    \script{GW150914_corner.py}
    \includegraphics[width=\columnwidth]{figures/GW150914_corner.png}
    \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
    The three sets of plots are the PEs done by LIGO with GR \citep{LIGO_2021},
    the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
    Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
    which does not suggest GW amplitude birefringence.}
    \label{fig:GW150914_corner}
\end{figure}

@@

\begin{enumerate}
    \item Introduction
    \begin{enumerate}
        \item Motivation
        \begin{enumerate}
            \item GR broke down at some scale, find a new theory
            \item CS: beyond-GR theory
            \item test CS property (GW amplitude birefringence)
        \end{enumerate}
        \item Amplitude Birefringence
        \begin{enumerate}
            \item GW amplitude birefringence is a property of space
            \item birefringence: enhance one and suppress another
            \item figure to illustrate the effect of birefringence
        \end{enumerate}
        \item Amplitude Ratio
        \begin{enumerate}
            \item GR: amplitude ratio depends on $\iota$ only
            \item nonGR: also depends on $\kappa$, $d_C$ and $f$
        \end{enumerate}
    \end{enumerate}
    \item Method
    \begin{enumerate}
        \item Parameter Estimation
        \begin{enumerate}
            \item PE on GWTC-3 to constrain $\kappa$
            \item Assume GW are generated as GR described
            \item birefringence is considered during the propagation
            \item waveform modification
        \end{enumerate}
    \end{enumerate}
    \item Results
    \begin{enumerate}
        \item GW150914
        \begin{enumerate}
            \item figure with frequency-dependent and frequency-independent
            \item posterior from $\cos\iota$ looks different
            \item frequency-dependent: degeneracy between $\iota$ and $\kappa$
            \item frequency-independent: degeneracy broken and $\kappa$ near $0$
        \end{enumerate}
        \item GWTC-3
        \begin{enumerate}
            \item figure of $\kappa$ of all events
            \item constraint on $\kappa$
            \item improvement of constraint on $\kappa$ compare to previous studies
        \end{enumerate}
    \end{enumerate}
    \item Discussion
    \begin{enumerate}
        \item Limitation
        \begin{enumerate}
            \item orientation of detector (polarisations)
            \item sensitivity of detector
        \end{enumerate}
        \item Future Study
        \begin{enumerate}
            \item More events from detector
            \item Higher order terms
        \end{enumerate}
    \end{enumerate}
\end{enumerate}

>>>>

\begin{figure}[h!]
    \script{birefringence.py}
    \includegraphics[width=\columnwidth]{figures/birefringence.png}
    \caption{
        The observed amplitudes of two polarisations of an artificial GW signal under different model.
        The top panel shows the case of GR, the amplitudes are unmodified.
        For the middle and bottom panel, they shows the case of birefringence.
        Left-handed polarisation is favored by the birefringence in the middle panel,
        and right-handed polarisation is favored by the birefringence in the bottom panel.
    }
    \label{fig:birefringence}
\end{figure}

GW amplitude birefringence is a property of space which consists of the enhancement of one GW polarisation over the other during the propagation of the wave.
The longer the GW travels, the larger the effect of birefringence.
To illustrate the effect of amplitude birefringence,
the observed amplitudes of two polarisations of an artificial GW signal example are shown in figure \ref{fig:birefringence}.
With birefringence, the amplitude is modified during the propagation,
and the observed amplitudes will be different compare to the amplitudes when they are generated at the binary.

@@

To constrain $\kappa$, we perform parameter estimation (PE) with data from the third LIGO-Virgo catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby,
a Bayesian toolkit for GW data analysis which is able to calculate posteriors of GW parameters based on interferometer data
and priors for the parameters in question. \citep{Bilby}

@@

After modifying Bilby, we set the prior on $\kappa$ to be a $\delta$ function at $0$, which is the same as using GR,
to test if the PEs could recover the PE done by LIGO in \citet{GWTC-2.1, GWTC-3}. These control tests result can be found in the data release.
We then set the prior on $\kappa$ to a uniform function between $-1$ and $1$, and perform PEs on the modified waveforms.
In both case, we use the same configuration as in \citep{GWTC-2.1, GWTC-3} for each individual events.

@@

\begin{figure}[h!]
    \script{GW150914_corner.py}
    \includegraphics[width=\columnwidth]{figures/GW150914_corner.png}
    \caption{
        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
    The three sets of plots are the PEs done by LIGO with GR \citep{GWTC-2.1, GWTC-3},
    the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
    Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
    which does not suggest GW amplitude birefringence.
    }
    \label{fig:GW150914_corner}
\end{figure}

@@

\begin{enumerate}
    \item Introduction
    \begin{enumerate}
        \item Motivation
        \begin{enumerate}
            \item GR broke down at some scale, find a new theory
            \item CS: beyond-GR theory
            \item test CS property (GW amplitude birefringence)
        \end{enumerate}
        \item Amplitude Birefringence
        \begin{enumerate}
            \item GW amplitude birefringence is a property of space
            \item birefringence: enhance one and suppress another
            \item figure to illustrate the effect of birefringence
        \end{enumerate}
        \item Amplitude Ratio
        \begin{enumerate}
            \item GR: amplitude ratio depends on $\iota$ only
            \item nonGR: also depends on $\kappa$, $d_C$ and $f$
        \end{enumerate}
    \end{enumerate}
    \item Method
    \begin{enumerate}
        \item Parameter Estimation
        \begin{enumerate}
            \item PE on GWTC-3 to constrain $\kappa$
            \item Assume GW are generated as GR described
            \item birefringence is considered during the propagation
            \item waveform modification
            \item GR control test
            \item same configuration as LIGO
        \end{enumerate}
    \end{enumerate}
    \item Results
    \begin{enumerate}
        \item GW150914
        \begin{enumerate}
            \item figure with frequency-dependent and frequency-independent
            \item posterior from $\cos\iota$ looks different
            \item frequency-dependent: degeneracy between $\iota$ and $\kappa$
            \item frequency-independent: degeneracy broken and $\kappa$ near $0$
        \end{enumerate}
        \item GWTC-3
        \begin{enumerate}
            \item figure of $\kappa$ of all events
            \item constraint on $\kappa$
            \item improvement of constraint on $\kappa$ compare to previous studies
        \end{enumerate}
    \end{enumerate}
    \item Discussion
    \begin{enumerate}
        \item Limitation
        \begin{enumerate}
            \item orientation of detector (polarisations)
            \item sensitivity of detector
        \end{enumerate}
        \item Future Study
        \begin{enumerate}
            \item More events from detector
            \item Higher order terms
        \end{enumerate}
    \end{enumerate}
\end{enumerate}
